Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-08-12 2_SSL_NE/article/article_supplementary.tex
added edits

\subsection{Luminous Efficacy of Radiation}
\label{subsec:ler}

Figure 1 in the main publication uses luminous efficacy of radiation as the primary metric to describe progress in lighting technologies. Care must be taken not to confuse this metric with luminous efficacy of source, which is used in \cref{fig:phosphor_spectrum}.

This metric describes the match of a light-emitting diode package spectrum to the human visual system. Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength-dependent human sensitivity to light \cite{cie-term-effrad}
%
\begin{myequation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{myequation}
%
where
%
\begin{align*}
    K &\dots \text{spectral luminous efficacy} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}
%
This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in the top panel of \cref{fig:phosphor_spectrum}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation. Care must be taken not to confuse this efficacy metric with the \textit{efficacy of source} described in the following subsection.

\subsection{Luminous Efficacy of Source}
\label{subsec:les}

The luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power \cite{cie-term-effsrc}
%
\begin{myequation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{myequation}
%
This metric is often cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. Care must be taken not to confuse this efficacy metric with the \textit{luminous efficacy of radiation}, which depends only on the spectral characteristics of a light source. As the luminous efficacy of a light source $\eta$ captures the overall device efficacy, it depends on a large number of other device properties and parameters. This makes attribution of changes in this metric to individual changes in device design or manufacturing difficult. For this reason, we do not use this metric in our study.

@@

@@

\clearpage
\subsection{LED Innovations}

\subsubsection{Non-Phosphor-Related}

\cref{tab:innovations} summarizes LED innovations and technology and manufacturing process improvements identified in this study as affecting key white LED sub-efficiencies. Note that phosphor-related LED innovations affecting consumer experience metrics, along with spectral efficiency, are presented separately in Table 1 in the main publication and discussed in detail in \cref{sec:innovation_phosphor}.

\begin{table}[h!]
    \caption{LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article.}
    \vspace{5mm}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \multicolumn{5}{c}{Light Conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{NiceTabularX}
    \vspace{5mm}

     Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.
    \label{tab:innovations}
\end{table}

\subsubsection{Phosphor-Related}
\label{sec:innovation_phosphor}

This section provides additional details for phosphor-related LED innovations identified in this study as affecting both spectral sub-efficiencies and consumer experience metrics.

\paragraph{YAG and YGAG Phosphors}

@@

\paragraph{258, SLA and SALON phosphors}

@@

\paragraph{PFS Phosphor}

@@

\paragraph{Quantum Dots for Light Down-Conversion}

@@

\clearpage
\paragraph{Spectral Data for Identified LED Phosphor Innovations}

\hspace{1mm}

@@

\newpage
\subsection{Cost Comparison with Reported Industry Data and DOE Projections}

We conducted a comparison of the LED manufacturing cost structure produced by our cost model with previously reported US DOE calculations and projections based on the LEDCOM model and industry data provided as part of industry round table discussions (see \cref{fig:costmodel_calibration}). We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model relying on state-of-the-art equipment at a virtual US-based manufacturing location, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In addition, the manufacturing lines of the majority of manufacturers is located in Asia. We also note that the share of the substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL}\cite{Xie2013}\cite{ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_calibration.pdf}
	\caption{Comparison of LED manufacturing cost structure between our manufacturing cost model (bottom panel) and previously published US DOE manufacturing cost calculations and projections based on industry data and LEDCOM model (top panel). Hatched bars are projections given by US DOE for 2018 and 2020. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in \cref{sec:costmodel}.}
	\label{fig:costmodel_calibration}
\end{figure}

@@

\begin{table}[h!]
    \caption{List of sources for Figure 1 organized by technology.}
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Technology} & \textit{References} \\
    \hline
    LED & own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{NiceTabularX}
    \caption{Note: CFL - Compact Fluorescent Lamp. HID - High-Intensity Discharge.}
\end{table}

>>>>

\subsubsection{Luminous Efficacy of Radiation}
\label{subsec:ler}

The luminous efficacy \textit{of radiation} $K$ describes the match of a light-emitting diode package spectrum to the human visual system. Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength-dependent human sensitivity to light \cite{cie-term-effrad}
%
\begin{myequation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{myequation}
%
where
%
\begin{align*}
    K &\dots \text{luminosity function} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}
%
This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in the top panel of \cref{fig:phosphor_spectrum}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation.

Care must be taken not to confuse this metric with luminous efficacy of source described in \cref{subsec:les}.

\subsubsection{Luminous Efficacy of Source}
\label{subsec:les}

The luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power $P_{el}$ \cite{cie-term-effsrc}. It is often used to describe general progress in lighting technologies.
%
\begin{myequation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{myequation}
%
This metric is also frequently cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. However, as noted in \cref{subsec:ler}, care must be taken not to confuse this efficacy metric with the \textit{luminous efficacy of radiation}, which depends only on the spectral characteristics of a light source. As the luminous efficacy of a light source $\eta$ captures the overall device efficacy, it depends on a large number of other device properties and parameters. This makes attribution of changes in this metric to individual changes in device design or manufacturing difficult. For this reason, we do not use this metric only to show historical progress across different lighting technologies in Figure 1 in the main text of the article.

@@

\clearpage
\section{Identified LED Innovations and Performance Improvements}

This Supplementary Note provides details for the identified LED innovations and performance improvements that contributed to the historical progress in LED technology across efficiency and consumer experience metrics discussed in the main text.

@@

\subsection{Non-Phosphor-Related LED Innovations}

\cref{tab:innovations} summarizes LED innovations and technology and manufacturing process improvements identified in this study as affecting four white LED sub-efficiencies: forward voltage efficiency, light extraction efficiency, internal quantum efficiency, droop and light conversion efficiency. Note that this table does not include phosphor-related LED innovations affecting spectral efficiency, as these are presented separately in Table 1 in the main text as affecting consumer experience metrics and discussed in detail in \cref{sec:innovation_phosphor}.

\begin{table}[h!]
    \caption{LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article.}
    \vspace{5mm}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
        \hline
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \multicolumn{5}{c}{Droop $\eta_{droop}$} \\
        \hline
            ongoing & Lower Current Density & Current Density & No & I \\
        \hline
            ongoing & Epitaxy \newline Improvements & $\ominus/\oplus$ Distribution \newline Defect Density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            ongoing & Chip Architecture \newline Improvements  & $\ominus/\oplus$ Distribution \newline Defect Density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect Getting \newline Underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \multicolumn{5}{c}{Light Conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect gettering \newline underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{NiceTabularX}
    \caption*{Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.}
    \label{tab:innovations}
\end{table}

\subsection{Phosphor-Related LED Innovations}
\label{sec:innovation_phosphor}

This section provides additional details for phosphor-related LED innovations identified in this study as affecting both spectral sub-efficiencies and consumer experience metrics. These include a detailed description of the history of those innovations, which we use to identify the details of involved technology spillovers. In addition, spectral data for the first identified LED products that used those innovations is presented in \cref{fig:phosphor_spectrum}, which we use to calculate consumer experience metrics, as well as LED spectral efficiency.

\subsubsection{YAG and YGAG Phosphors}

@@

\subsubsection{258, SLA and SALON phosphors}

@@

\subsubsection{PFS Phosphor}

@@

\subsubsection{Quantum Dots for Light Down-Conversion}

@@

\clearpage
\subsubsection{Spectral Data for Identified LED Phosphor Innovations}

@@

\newpage
\section{Cost Model Sensitivity Analysis}

\begin{figure}[ht!]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_sensitivity.pdf}
	\caption{Sensitivity analysis for selected cost model parameters. Panel a represents the tested range of variation of parameters indicated on horizontal axis along with their respective measurement units, as represented in Supplementary Table 3. Panel b shows the resulting range of variation in the manufacturing cost. Ranges are represented for three years considered in the cost model, where corresponding data is available, by whiskers of the following colour: 2003 – blue, 2012 – green, 2020 – red. Note the overall trend of decreasing sensitivity to cost model parameters as the number of die per wafer increases over time. Abbreviations: resp. - respective; Yield Sav. from Insp. - yield savings from inspection.}
	\label{fig:sensitivity}
\end{figure}

A sensitivity analysis of the cost model has been performed using parameter variations listed in table \cref{tab:sensitivity}. The amount of parameter variation was chosen to encompass the identified range of parameter values in different manufacturing setups for each considered year. As an example, the lower range in the variation for the price of metals used in the model (+50\%,-0\%) was chosen because the prices quotes from the United State Geological Survey price database \cite{usgeoprices}. Industry metals are not sold below the price of the raw material and often the markup is small compared to the price of the material. The upper range was chosen because a survey of industrial metal suppliers for semiconductor manufacturing showed that the largest markup was below 50\%. The results of the analysis are shown in  \cref{fig:sensitivity}. The cost model is generally more sensitive to the variation in parameters at smaller wafer diameters. The most sensitive parameters are global parameters, such as yield or average equipment throughput.

\begin{table}[]
\small
    \caption{Cost model sensitivity analysis parameter list. The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}.}
    \vspace{5mm}
    \begin{NiceTabularX}{\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Person-hour & FTE & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 100\% & +25,-25 & 100\% & +25,-15 & 100\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge Exclusion & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & 100\% & +50,-0 & 100\% & +50,-0 & 100\% & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & 100\% & +50,-50 & 150 & +0,-0 & 200 & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{NiceTabularX}
    \caption*{Note: The units for the values in columns \textit{2003}-\textit{2020} are indicated in the column \textit{Units}. If values in columns \textit{2003}-\textit{2020} are instead given in \%, this indicates that the parameters were varied by a set percentage from their respective model baselines. Abbreviations: FTE - full-time equivalent; UPH - units per hour; Equip. Discount - equipment discount (sales rebate for large purchases); Inspec. Yield Savings - yield savings from inspection (early detection and alleviation of issues in the manufacturing workflow).}
    \label{tab:sensitivity}
\end{table}

\clearpage
\section{Cost Comparison with Reported Industry Data and DOE Projections}

We conduct a comparison of the outcomes of our LED manufacturing cost model regarding the evolution of the cost structure of low-to-mid-power classic chip GaN phosphor-converted white LED with previously reported U.S. DOE calculations and projections based on the LEDCOM model and industry data provided as part of industry round table discussions (see \cref{fig:costmodel_calibration}). We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model relying on state-of-the-art equipment at a virtual U.S.-based manufacturing location, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In addition, the manufacturing lines of the majority of manufacturers are located in Asia. We also note that the share of the sapphire substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections made for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL,Xie2013,ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections.

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_calibration.pdf}
	\caption{Top panel: manufacturing cost calculations (2008-2016) and projections (2018, 2020) previously published by the U.S. DOE based on industry data and LEDCOM model. Bottom panel: manufacturing cost structure produced by our cost model for 2002, 2012 and 2020. Hatched bars are DOE projections. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in \cref{sec:costmodel}.}
	\label{fig:costmodel_calibration}
\end{figure}

@@

\begin{table}[h!]
    \caption{List of sources for Figure 1 organized by technology.}
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Technology} & \textit{References} \\
    \hline
    LED & own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{NiceTabularX}
    \caption*{Note: CFL - Compact Fluorescent Lamp. HID - High-Intensity Discharge.}
\end{table}
